How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985

### 2022-09-09 Draft_September_2022/Full_draft_20220901.rmd
#implementedgrammarcorrectionsfromAnna #polishedfigure4 #correctedafewplotannotations

---
title: How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "7 September 2022"
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    ```{=latex}
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
    ```
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about, and act towards, the environment transnationally. Still, these remain understudied in this context. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses along time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency of the Amazon as a topic in presidential speeches inside and outside Brazil. While constructing the Amazon a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest presidents are from the Amazon, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

@@

  mutate(mx_cat = factor(case_when(mixed_type == '1_0_0_0' ~ "Pure Sovereignty",

@@

                                   mixed_type == '0_0_0_1' ~ "Pure Conservation",

@@

         levels = c("Pure Economic Integration", "Pure Conservation",

@@

                    "Pure Sovereignty", "Other", "Economic Conservation", 

@@

         mx_cat2 = case_when(grepl("Pure Economic Integration|Pure Conservation|
                                   |Pure Social Development|Pure Sovereignty",

@@

We need to protect the Amazon from foreign interests. We need to exploit the Amazon’s natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem. Each of these statements contains an implicit assumption of what needs to be solved, or in other words, it represents the Amazon as a particular problem. These specific problems touch on shared imaginaries of the Amazon as part of the larger socio-cultural history of the country, for example, in relation to issues of national sovereignty, economic integration, social development, or environmental conservation.

Different governments throughout Brazilian history have been described as proponents of a specific view of the Amazon [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. We conceptualize the Amazon region, forest, and people to be a policy object, this is a specific issue that deserves dedicated policy attention. While the military dictatorship is associated with understanding the Amazona as issues of national sovereignty and economic integration, Sarney’s and Lula’s presidencies are often tied to environmental conservation views. These monolithic representations advance that specific governments understand and act towards the Amazon as one problem, hiding within government diversity of both policies and perspectives. Albeit the current calls to understand the environment as a social-cultural construction and to identify the effect of culture on environmental outcomes [@lepolaindewaroux2021], we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between, or within, governments. Building on @hirschman1963 conceptualization of chosen problems and @bacchi2009 theory of problem-representation in policy, we propose a framework for identifying problem-constructions in discourse and investigate how the Brazilian Amazon has been constructed as a problem in transnational political discourses. We understand transnationalism^[ We use the noun transnationalism, the adverb transnationally, and the adjective transnational interchangeably] as relationships that transcend nation-states, from the local level to the supranational level encompassing non-state actors [@keck1998]

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022a], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015], and many others. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices [@assuncao2015; @capobianco2019; @capobianco2021], are a crucial factor in decisions to deforest on the ground [@campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different sites, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

To investigate how the Brazilian Amazon has been constructed as a problem in presidential speeches, we create a dataset containing 6130 official presidential speeches by all Brazilian presidents since 1985. We subset the dataset by identifying Amazonian-related statements within these speeches. We find that 2014 sections in these discourses refer to the Amazon at least once. We then develop a codebook grounded on Amazonian historiography to code how each of these statements constructs Amazon as a particular problem. We use this codebook to manually code a randomly selected training set of Amazonian-related statements. We train a supervised machine-learning model to automatically label the remaining set of Amazonian statements. We then conduct a descriptive and inferential analysis of this data, tying patterns in the data to Amazonian policies and deforestation outcomes over time.

Our empirical findings are threefold. First, we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (UNFCCC Summits). Second, the same government adopts a diverse range of problem-constructions over time. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, briefly surpassing economic integration as more common problem-constructions from 2011 to 2016. Constructing the Amazon as an issue of sovereignty has increasingly become more pertinent after 2010. Finally, using logit models, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation as presidents move away from the Amazon region. That is, presidents, talk to the economic and social needs of the people domestically while boasting about environmental policy internationally.

@@

In “Journey towards Progress”, @hirschman1963 draws a conceptual distinction between pressing problems (pressured by outside parties to the government) and chosen problems (chosen by the government at their discretion). Pressing problems can be either privileged or neglected, depending on the degree of pressure exercised by the interested groups. Chosen problems are those governments select at their discretion. Problems can change from pressing to chosen across time and in space as a function of (a) solutions becoming available, (b) a change in the level of government control in society, or (c) a shift of interests from top policymakers [@hirschman1963]. The author exemplifies chosen problems with the case of the construction of Brasilia.

@bacchi2009 [p. 10] argues that policies have a cultural dimension as “it takes shape within specific historical and national or international contexts”. The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed, this is a policy object. We conceptualize policy objects as a problem that demands dedicated policy attention. The alleged problem is not always explicitly stated in policy: policies are represented as solutions for implicit problems. Building Brasilia (the policy object), thus, could solve a problem of regional inequality, a problem of a dormant economy without state investment, a problem of political representation, or all three depending on how governments speak about it.

Governments can choose to emphasize (or not) one or more implicit problems that a policy solves depending on which interest group they are in communication with. Relative to how the policy is represented to be, it can be a solution to problems that are considered pressing or not for different groups and it is up to the discretion of the political actor to construct a particular problem in a particular way given context. @putnam1988 seminal article on the two-level game can help us make sense of this variation conceptually. The author argues that the outcomes of international negotiations lie within the overlap between the agenda (and pressure) of domestic groups and international groups. If policies are socially constructed, and consequently their implicit problem can vary, governments can in theory construct them to meet the expectations of both. Consequently, this entails that governments are more diverse in their opinions, rather than monolithic.

An interesting site to study the relationship between problem-construction and policy objects is presidential speeches, as they legitimize ways of thinking about an issue. While the relationship between presidential speeches and policy may not be causal, they have the power to introduce and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004; @gillion2016]. As presidents speak across several locations to various audiences, they frame similar issues differently.

@@

The connection between presidential discourse and the environment has been studied in the case of the United States but remains conceptually underdeveloped. [@calderwood2019], for instance, examines 2019 mentions of climate change in American official presidential speeches since 1989. Among other things, the author finds that American presidents frequently side-step the environmental aspects of climate change, even as frames move from focusing on economics to focusing on security aspects of climate change. Building on @putnam1988, @calderwood2020 demonstrates that American presidents are more likely to mention climate change in foreign locations, and that location influences the specific discursive approach and tone they adopt. Elsewhere, @bevitori2015 finds that mentions of the environment are typically co-selected with the pronoun ‘our’, as well as with ‘economy’, ‘clean’, and ‘preserve’. While these studies corroborate the possibilities of variation depending on the audience, they fail to tie vocabulary to specific problem-constructions based on wider shared meanings of environment and climate change. That is, problem-constructions touch on shared meanings available to both the speaker and the audiences as part of larger social-cultural history [see @bacchi2009], therefore, mentions to the lexicon as ‘environment’, ‘preserve’, or ‘climate change should be tied to the larger meaning of policies in the United States.

There is also an empirical gap in terms of how the Amazon, specifically, has been constructed as a problem along time and geographic location in political discourses. Whereas governmental discourses in Brazil have been studied for topics such as inflation or race relations [see @silva2019], the environment remains absent. An exception is @barros2020, who investigates Amazonian discourse in the Brazilian Congress. The author finds that the economic value of the Amazon for the cattle industry is the most salient narrative put forward by congressmen and concludes that there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

This variation in problem-construction entails that the same governments are more diverse in their positions than the literature suggests. Scholars often describe federal governments as proponents of a cohesive set of policies toward the Amazon [see @drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. In this literature, the 1964 military dictatorship is associated with securing sovereignty in the region by populating it and integrating it into the national economy [@drummond2006]. The governments from the late 1980s up to 2009 are associated with a turn towards policies that focus on environmental conservation. The presidencies of Dilma and Temer are connected to the decline of conservationist policies, while Bolsonaro to the dismantling of environmental policies. Whereas these works are important to understand how different governments acted toward the Amazon, the classification of specific governments into policy periods, or policy cycles, represents them as monolithic: they associate specific governments with one specific view of the Amazon.

We propose a conceptual framework that accepts the possibility of varied problem-construction for the same policy object and connects it to presidential speeches. Governments often choose what problems to solve and what policies to implement. The same policies can be represented as solving different problems implying a degree of social construction [@bacchi2009]. Governments can emphasize, or not, implicit problems that a policy solves depending on which interest group they are in communication with. The variation in problem-construction suggests that Brazilian governments are more diverse in their positions than the literature suggests. An interesting site to study the relationship between varied problem-construction and policy is presidential speeches as they legitimize ways of thinking about an issue, matter for policymaking, and environmental outcomes. As presidents speak in different places and to diverse audiences, it is an empirical site prone to identifying variation.

@@

The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon. In 1966, the Brazilian Military launched Operation Amazon, a policy to modernize the region based on three assumptions @acker2021. First, it assumes that nature should be conquered by men. Second, it assumes that exploiting natural resources would render the Amazon region economically profitable. Third, it assumes that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructure projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. @capobianco2019 describes the period from the 1950s to the 1980s in relation to the wider range of policies of economic integration, often centralized by the federal government [see @becker2005]. The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region into the national and international economy.

@@

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. A common explanation for the creation of these institutions in the Amazonian literature is the impression of the lack of control over the market engendered years of centralized economic integration in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. @hochstetler2007 argue that during preparations for the summit, a new form of Brazilian environmentalism emerged: socio-environmentalism. They define it as an emphasis on the local livelihoods of traditional populations while protecting nature. @capobianco2019 argues that socio-environmentalism informed a series of policies throughout the 1990s and early 2000s towards preserving the forest, such as the establishment of several conservation units in the Amazon in 2001. The environmental conservation problem-constructions emphasize the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

@@

Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples’ dignity, standards of living, and other constitutional rights. However, such emphasis can be masked within both the economic integration and environmental conservation accounts. @acker2021 [p. 10], for example, argues that during the military dictatorship attempts to populate the Amazon was seen, among other things, as a solution for the country’s growing social inequalities elsewhere. As democracy established itself in Brazil, the settlers from the military dictatorship became electorates with constitutional rights. @hochstetler2007 argue that with the transition to democracy, the mobilization of diverse social movements to demand social equality and conservation was an integral part of the socio-environmentalism movement in Brazil. Whereas indigenous rights and preservation of the forest go hand in hand, it is not clear that providing better standards of livings to the large and poor populations of the state and preservation do; perhaps this is also why there was never an open consensus about socio-environmentalism in Brazil [@hochstetler2007]. While related to economic integration and environmental conservation, social development is a problem-construction that focuses on constitutional rights, citizenship, and dignity. The policy solution to social development implies the policies to facilitate access to water, sanitation, electricity, internet, and radio, as well as the construction of schools and hospitals locally.

@@

To analyze how presidents construct the Amazon as a problem, we build upon the dataset provided by [@cezar2020] which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil. Then, we proceed to identify all speeches that refer to the Amazon as a region, people, or forest in these speeches. We do so by detecting all speeches in which the stem “amazon” appears. In Portuguese, the stem captures terms such as “Amazonia”, “Amazonica”, “Amazonidas”, “Amazonense(s)”, “Amazonas”, among others. We find that 946 speeches are, at least partially, about the Amazon from the 6130.

Using the poldis R package [@sposito2021], we proceed to extract two sentences before and two sentences after the sentence in which the stem “amazon” appears. We opt for picking two sentences before and two sentences after, rather than words because sentences usually contain a cohesive idea. By doing so we create our unit of analysis: an Amazonian statement. We use Amazon statements as our unit of analysis for two main reasons. First, working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. Second, it increases the number of our observations and their variety meaningfully, allowing for more specificity in our analysis. This process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified. When an Amazonian statement contains two or more matches of the stem “amazon”, we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

@@

Second, with the codebook in hand, each one of the authors, separately, hand-coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazon Statements identified. We chose to hand code half of the observations because since there are several nuances in discourse in how presidents talk about the amazon in time and as a problem, we conservatively code a relatively large training set for the subsequent automated machine learning. As well, this allows a robust validation set to verify the models. As the size of the training set should increase with the number of categories [@grimmer2022], we deemed four categories and 50% of the statements as reliable. Also, automating the coding of half of the observations saved the authors over one month of work in comparison to manual coding. Intercoder agreement for each of the four main categories was 85%, on average. For each non-matching coded observation, the co-authors discussed and sorted their disagreements.

@@

When, where, and how often do presidents speak about the Amazon? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents mention the Amazon in 15% ^[For comparison purposes the averages for other policy objects in the same corpus of speeches are: Inequality (14.7%), Criminality (17.3%), Inflation (20.9%), and Unemployment (13.8%).] of their speeches. The frequency in which the Amazon appears across presidential speeches inside, outside, and overall varies similarly across time (see Figure 7 in appendix).

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "9 September 2022"
output:
  html_document:
    df_print: paged
  latex_engine: xelatex
  fig_caption: yes
  pdf_document: default
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about and act towards the environment transnationally, but they are understudied. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses along time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency at which the topic of the Amazon is mentioned in presidential speeches inside and outside Brazil. While constructing the Amazon as a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest away presidents are from the Amazon itself, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

@@

  mutate(mx_cat = factor(case_when(mixed_type == '1_0_0_0' ~ "Pure National Sovereignty",

@@

                                   mixed_type == '0_0_0_1' ~ "Pure Environmental Conservation",

@@

         levels = c("Pure Economic Integration", "Pure Environmental Conservation",

@@

                    "Pure National Sovereignty", "Other", "Economic Conservation", 

@@

         mx_cat2 = case_when(grepl("Pure Economic Integration|Pure Environmental Conservation|
                                   |Pure Social Development|Pure National Sovereignty",

@@

*We need to protect the Amazon from foreign interests. We need to exploit the Amazon’s natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem.* Each of these statements contains an implicit assumption of what needs to be solved, or in other words, it represents the Amazon as a particular problem. These specific problems touch on shared imaginaries of the Amazon as part of the larger socio-cultural history of the country, for example, in relation to issues of national sovereignty, economic integration, social development, or environmental conservation.

Different governments throughout Brazilian history have been described as proponents of a specific view of the Amazon [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. We conceptualize the Amazon region, forest, and people to be a policy object, this is a specific issue that deserves dedicated policy attention. While the military dictatorship is associated with understanding the Amazon as issues of national sovereignty and economic integration, Sarney’s and Lula’s presidencies are often tied to environmental conservation views. These monolithic representations advance that specific governments understand and act towards the Amazon as one problem, hiding within government diversity of both policies and perspectives. Albeit the current calls to understand the environment as a social-cultural construction and to identify the effect of culture on environmental outcomes [@lepolaindewaroux2021], we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between, or within, governments. Building on @hirschman1963 conceptualization of chosen problems and @bacchi2009 theory of problem-representation in policy, we propose a framework for identifying problem-constructions in discourse and investigate how the Brazilian Amazon has been constructed as a problem in transnational political discourses. We understand transnationalism^[ We use the noun transnationalism, the adverb transnationally, and the adjective transnational interchangeably] as relationships that transcend nation-states, from the local level to the supranational level encompassing non-state actors [@keck1998]

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022a], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015], and many others. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices [@assuncao2015; @capobianco2019; @capobianco2021], are a crucial factor in decisions to deforest on the ground [@campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different settings, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

To investigate how the Brazilian Amazon has been constructed as a problem in presidential speeches, we create a dataset containing 6130 official presidential speeches by all Brazilian presidents since 1985. We subset the dataset by identifying Amazon-related statements within these speeches. We find that 2014 sections in these discourses refer to the Amazon at least once. We then develop a codebook grounded on Amazonian historiography to code how each of these statements constructs Amazon as a particular problem. We use this codebook to manually code a randomly selected training set of Amazon-related statements. We train a supervised machine-learning model to automatically label the remaining set of Amazonian statements. We then conduct a descriptive and inferential analysis of this data, tying patterns in the data to Amazonian policies and deforestation outcomes over time.

Our empirical findings are threefold. First, we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (Climate Summits). Second, the same government adopts a diverse range of problem-constructions over time. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, briefly surpassing economic integration as more common problem-constructions from 2011 to 2016. Constructing the Amazon as an issue of sovereignty has increasingly become more pertinent after 2010. Finally, using logit models, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation as presidents move farther away from the Amazon region. That is, presidents talk to the economic and social needs of the people domestically while boasting about environmental policy internationally. Conceptually, this contributes to understanding the social construction of the Amazon in discourses and how these relates to policies. Empirically, we provide the first comprehensive overview of when, where, and how the Amazon is constructed as problems in presidential speeches.

@@

In “Journey towards Progress”, @hirschman1963 draws a conceptual distinction between pressing problems (pressured by outside parties to the government) and chosen problems (chosen by the government at their discretion). Pressing problems can be either privileged or neglected, depending on the degree of pressure exercised by the interested groups. Chosen problems are those that governments select at their discretion. Problems can change from pressing to chosen across time and in space as a function of (a) solutions becoming available, (b) a change in the level of government control in society, or (c) a shift of interests from top policymakers [@hirschman1963]. Hirschman exemplifies chosen problems with the case of the construction of Brasilia.

@bacchi2009 [p. 10] argues that policies have a cultural dimension as they take“ shape within specific historical and national or international contexts”. The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed, this is a policy object. We conceptualize policy objects as a problems that demand dedicated policy attention. The alleged problem is not always explicitly stated in a policy: policies are represented as solutions for implicit problems. Building Brasilia (the policy object), thus, could solve a problem of regional inequality, a problem of a dormant economy without state investment, a problem of political representation, or all three depending on how governments speak about it.

Governments can choose to emphasize (or not) one or more implicit problems that a policy solves depending on which interest group they are in communication with. Depending on how the policy is represented to be, it can be portrayed as a solution to problems that are considered pressing or not for different groups and it is up to the discretion of the political actor to construct a particular problem in a particular way given context. @putnam1988's seminal article on the two-level game can help us make sense of this variation conceptually. The author argues that the outcomes of international negotiations lie within the overlap between the agenda (and pressure) of domestic and international groups. If policies are socially constructed, governments can in theory construct them to meet the expectations of both. Consequently, this entails that governments are more diverse in their opinions, rather than monolithic.

An interesting context to study the relationship between problem-construction and policy objects is presidential speeches, as they legitimize ways of thinking about an issue. While the relationship between presidential speeches and policy may not be causal, they have the power to introduce and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004; @gillion2016]. As presidents speak across several locations to various audiences, they frame similar issues differently.

@@

The connection between presidential discourse and the environment has been studied in the case of the United States but remains conceptually underdeveloped.@calderwood2019, for instance, examines 2019 mentions of climate change in American official presidential speeches since 1989. Among other things, the author finds that American presidents frequently side-step the environmental aspects of climate change. Building on @putnam1988, @calderwood2020 demonstrates that American presidents are more likely to mention climate change in foreign locations, and that location influences the specific discursive approach and tone they adopt. Elsewhere, @bevitori2015 finds that mentions of the environment are typically co-selected with the pronoun ‘our’, as well as with ‘economy’, ‘clean’, and ‘preserve’. While these studies corroborate the possibilities of variation depending on the audience, they fail to tie vocabulary to specific problem-constructions based on wider shared meanings of environment and climate change. That is, problem-constructions touch on shared meanings available to both the speaker and the audiences as part of larger social-cultural history [see @bacchi2009], therefore, mentions to the lexicon as ‘environment’, ‘preserve’, or ‘climate change should be tied to the larger meaning of policies in the United States.

There is also an empirical gap in terms of how the Amazon, specifically, has been constructed as a problem along time and geographic location in political discourses. While governmental discourses in Brazil have been studied for topics such as inflation or race relations [see @silva2019], the environment remains largely absent. An exception is @barros2020, who investigates Amazonian discourse in the Brazilian Congress. The author finds that the economic value of the Amazon for the cattle industry is the most salient narrative put forward by congressmen and concludes that there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

This variation in problem-construction entails that the same governments are more diverse in their positions than the literature suggests. Scholars often describe federal governments as proponents of a cohesive set of policies toward the Amazon [see @drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. In this literature, the 1964 military dictatorship is associated with securing sovereignty in the region by populating it and integrating it into the national economy [@drummond2006]. The governments from the late 1980s up to 2009 are associated with a turn towards policies that focus on environmental conservation. The presidencies of Dilma and Temer are connected to the decline of conservationist policies, while Bolsonaro to the dismantling of environmental policies. Although these works are important to understand how different governments have acted toward the Amazon, the classification of specific governments into policy periods, or policy cycles, represents them as monolithic: they associate specific governments with one specific view of the Amazon.

We propose a conceptual framework that accepts the possibility of varied problem-construction for the same policy object and connects it to presidential speeches. Governments often choose what problems to solve and what policies to implement. The same policies can be represented as solving different problems implying a degree of social construction [@bacchi2009]. Governments can emphasize, or not, implicit problems that a policy solves depending on which interest group they are in communication with. The variation in problem-construction suggests that Brazilian governments are more diverse in their positions than the literature suggests. An interesting context to study the relationship between varied problem-construction and policy is presidential speeches as they legitimize ways of thinking about an issue, matter for policymaking, and environmental outcomes. As presidents speak in different places and to diverse audiences, it is an empirical site prone to identifying variation.

@@

The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon. In 1966, the Brazilian Military launched 'Operation Amazon', a policy to modernize the region based on three assumptions @acker2021. First, it assumes that nature should be conquered by mankind Second, it assumes that exploiting natural resources would render the Amazon region economically profitable. Third, it assumes that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructural projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. @capobianco2019 and @becker2005 describe the period from the 1950s to the 1980s in similar vein, as one of often centralized economic policies by the federal government.The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region into the national and international economy.

@@

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. A common explanation for the creation of these institutions in the Amazonian literature is the impression of the lack of control over the market engendered years of centralized economic integration in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. @hochstetler2007 argue that during preparations for the summit, a new form of Brazilian environmentalism emerged: socio-environmentalism. They define it as an emphasis on the local livelihoods of traditional populations while protecting nature. @capobianco2019 argues that socio-environmentalism informed a series of policies throughout the 1990s and early 2000s towards preserving the forest, such as the establishment of several conservation units in the Amazon in 2001. The environmental conservation problem-constructions emphasize that the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

@@

Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples’ dignity, standards of living, and other constitutional rights. However, such emphasis can be masked within both the economic integration and environmental conservation accounts. @acker2021 [p. 10], for example, argues that during the military dictatorship attempts to populate the Amazon were seen, among other things, as a solution for the country’s growing social inequalities elsewhere. As democracy established itself in Brazil, the settlers from the military dictatorship became electorates with constitutional rights. @hochstetler2007 argue that with the transition to democracy, the mobilization of diverse social movements to demand social equality and conservation was an integral part of the socio-environmentalist movement in Brazil. While indigenous rights and preservation of the forest go hand in hand, it is not clear whether providing better standards of livings to the large and poor populations of the state and preservation do; perhaps this is also why there was never an open consensus about socio-environmentalism in Brazil [@hochstetler2007]. Although related to economic integration and environmental conservation, social development is a problem-construction that focuses on constitutional rights, citizenship, and dignity. The policy solution to social development implies the policies to facilitate access to water, sanitation, electricity, internet, and radio, as well as the construction of schools and hospitals locally.

@@

To analyze how presidents construct the Amazon as a problem, we build upon the dataset provided by [@cezar2020] which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil. Then, we proceed to identify all speeches that refer to the Amazon as a region, people, or forest in these speeches. We do so by detecting all speeches in which the stem “amazon” appears. In Portuguese, the stem captures terms such as “Amazonia”, “Amazonica”, “Amazonidas”, “Amazonense(s)”, “Amazonas”, among others. We find that 946 speeches are, at least partially, about the Amazon out of the 6130.

Using the poldis R package [@sposito2021], we proceed to extract two sentences before and two sentences after the sentence in which the stem “amazon” appears. We opt for picking two sentences before and two sentences after, rather than words because sentences usually contain a cohesive idea. By doing so we create our unit of analysis: an Amazon statement. We use Amazon statements as our unit of analysis for two main reasons. First, working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. Second, it increases the number of observations and their variety meaningfully, allowing for more specificity in the analysis. This process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified. When an Amazonian statement contains two or more matches of the stem “amazon”, we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

@@

Second, with the codebook in hand, each one of the authors, separately, hand-coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazon Statements identified. We chose to hand.code half of the observations because since there are several nuances in discourse in how presidents talk about the Amazon over time and as a problem, we conservatively coded a relatively large training set for the subsequent automated machine learning. As well, this allows a robust validation set to verify the models. As the size of the training set should increase with the number of categories [@grimmer2022], we deemed four categories and 50% of the statements as reliable. Also, automating the coding of half of the observations saved the authors over one month of work in comparison to manual coding. The intercoder agreement for each of the four main categories was 85%, on average. For each non-matching coded observation, the co-authors discussed and sorted their disagreements.

@@

When, where, and how often do presidents speak about the Amazon? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents mention the Amazon in 15% ^[For comparison purposes the averages for other policy objects in the same corpus of speeches are: Inequality (14.7%), Criminality (17.3%), Inflation (20.9%), and Unemployment (13.8%).] of their speeches. The frequency at which the Amazon appears across presidential speeches inside, outside, and overall varies similarly across time (see Figure 7 in appendix).
